---
title: "Michael Hornsby's EDRM 718 Final Project Functions script"
output: html_notebook
---

Import the usual libraries that will be used for this project

```{r include=FALSE}
library(here)
library(tidyverse)
library(ggrepel)
library(scales)
library(readxl)
library(purrr)
library(dplyr)
library(car)
```


This is the function to highlight large residuals.  The function will model y
as a function of x and return a logical indicating if each x,y pair is 
one of the p largest.

```{r}
resid.function <- function(xvar, yvar, p, key) {
  
# Create the model and residuals
  
model <- lm(yvar ~ xvar)

residuals <- residuals(model)

my_data <- as.data.frame(cbind(key, xvar, yvar, residuals))

# Create an order vector and order the residuals from largest to smallest
order_vector <- order(abs(my_data$residuals), decreasing = TRUE)

my_data <- my_data[order_vector,]

# Create the logical vector, add it to my data set, and return the data set

logical_vector <- (c(rep(TRUE, p), rep(FALSE,length(my_data$xvar) - p)))

my_data$outcome <- logical_vector

my_data <<- data.frame(my_data)

# This creates the scatterplot with the regline.

scatterplot(x = my_data$xvar,
            y = my_data$residuals,
            regLine = FALSE,
            groups = my_data$outcome,
            legend = FALSE,
            smooth = FALSE,
            xlab = "Semester 1 GPA",
            ylab = "Semester 2 GPA",
            main = "Highlighted Largest Residuals")

return(my_data)
}
```


This is the function to reverse score items.
Numeric is the number of choices on the Likert Scale for the items.
Reverse is a vector of the items numbers that are reverse scored.

```{r}
reverse.function <- function(dataset, numeric, reverse) {

reverse <- reverse + 1
numeric <- numeric + 1
reverse_dataset <- dataset

for (i in reverse) {
  reverse_dataset[i] <- abs(reverse_dataset[i] - numeric)
}

dataset$total <- rowSums(reverse_dataset[,2:as.numeric(ncol(reverse_dataset))])

print(dataset)

dataset.tot <<- dataset

}
```

