---
title: "Michael Hornsby's EDRM 718 Final Project Script"
output: html_notebook
---

```{r include=FALSE}
library(here)
library(tidyverse)
library(ggrepel)
library(scales)
library(readxl)
library(purrr)
library(dplyr)
library(car)
```


Import the three data sets.  (Note: tidying was completed in Excel previously)

```{r echo=FALSE}
demo <- as.tibble(read_csv(here("data", "job_prep_demo.csv")))

demo

GPA <- as.tibble(read_csv(here("data", "job_prep_GPA.csv")))

GPA

survey <- as.tibble(read_csv(here("data", "job_prep_survey.csv")))

survey
```


Let's check the ID keys.  I want to make sure there is no ID that appears more
than once in the three datasets.

```{r}
check.demo <- demo %>% count(ID)
check.demo %>% filter(n > 1)
```


```{r}
check.survey <- survey %>% count(ID)
check.survey %>% filter(n > 1)
```


```{r}
check.GPA <- GPA %>% count(ID)
check.GPA %>% filter(n > 1)
```

We see that ID 659 appears twice in the GPA dataset. Let's look
at these observations.

```{r}
GPA %>% filter(ID == 659)
```

If the ID was simply a duplicate observation I would only remove one of them.
However, because the two are different, I'm going to remove both to be safe.

```{r}
GPA <- GPA %>% filter(ID != 659)

check.GPA <- GPA %>% count(ID)
check.GPA %>% filter(n > 1)
```

I removed the two observations where the ID was 659.  The number of 
observations in GPA dropped from 739 to 737, and I ran the check
again just to make sure there were no duplicate ID's.

Let's move to examining student GPAs.

